The binary liquid model for gravity-driven microchannel simulations

### 2010-10-08 Text/paper.tex
Modified sections about dimensionalization and initialization. Need to come out with nice tests.

In terms of D2Q9 the equilibrium populations are as \cite{pooley-contact}:
\begin{equation}
\begin{aligned}
&f_i^{eq}=w_i 
\biggl(3
p_0 - k \phi \Delta \phi
+\frac{u_{\alpha}c_{i\alpha}}{c_s^2}+\frac{Q_{i\alpha\beta}u_{\alpha } u_ {
\beta}}{2 c_s^4}\biggr), i=1\div8\\
&f_0^{eq}=\rho-\sum_{i\neq0}{f_i^{eq}}\\
&g_i^{eq}=w_i(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi^m
\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}), i=1\div8 \\
&g_0^{eq}=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
\end{aligned}
\end{equation}
where $\Gamma$ is the mobility parameter; chemical potential
$\mu=-A\phi+A\phi^3-k\Delta\phi$; $k$ is related to the surface
tension; $A$ is the parameter of the free-energy model; bulk pressure
$p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$. D2Q9 related parameters are
$w_i=\left\{\frac{4}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},
\frac{1}{36},\frac{1}{36},\frac{1}{36},\frac{1}{36}\right\}$, and
$Q_{i\alpha\beta}=c_{i\alpha} c_{i\beta} - c_s^2 \delta_{\alpha\beta}$ with
the sound speed parameter $c_s^2=\dfrac{1}{3}$. Weights related to the
inclusion of the surface tension coefficient are as follows:
$w^{xx}_{1-2}=w^{yy}_{3-4}=1/3$, $w^{xx}_{3-4}=w^{yy}_{1-2}=-1/6$,
$w^{xx}_{5-8}=w^{yy}_{5-8}=-1/24$, $w^{xy}_{1-4}=0$, $w^{xy}_{5-6}=1/4$ and
$w^{xy}_{7-8}=-1/4$. Note that the system restores the macroscopic
equations as:
\begin{equation}
\begin{aligned}
&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}=
-\partial_{\alpha}P_{\alpha \beta} +
\nu\partial_{\beta}\left(\partial_{\alpha}u_{\beta}+\partial_{\beta} u_{\alpha}
+ \frac{1}{3}\partial_{\gamma} u_{\gamma} \delta_{\alpha\beta}\right)\\
&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \Delta \mu,
\end{aligned}
\label{binary:fluid:system}
\end{equation}
where the viscosity $\nu=c_s^2 (\tau-1/2)$ and mobility parameter
$M=\Gamma(\tau_{\phi}-1/2)$. The system allows the separation of the liquid
phase with $\phi=1$ and so-called gas phase with $\phi=-1$. Note that the
spatial variation of the relaxation parameter depends on the relaxation
parameter of $\tau_{gas}$ and $\tau_{liq}$ linearly as
$\tau=\tau_{gas}+\frac{\phi+1}{2}\tau_{liq}$.

\section{Lattice Botlzmann benchmark}
To properly simulate and to compare simulations results with the already
performed simulations one needs to design a benchmark by addressing the
following challenges for the lattice Boltzmann method:
\begin{itemize}
 \item The diffusive interface nature of the multiphase model should not
influence the film resolution.
 \item The spurious currents arising from the surface tension discretization
can influence the mass transfer in the thin film therefore comprimising the
overall solution.
 \item The wettability coefficient which is in principle can control the
dynamic angle and menisci appearance \cite{pagonabarraga-finger} should not
influence on the overall solution.
 \item The free-energy model examined in the work is one density/different
viscosities model. Classical Bretherton problem is the free-interface problem
where the film thickness is established by itself. The viscosities ratio should
be large as it is indicated in \cite{pagonabarraga-fingers}, however no studies
address this issue. Because the free-interface nature of the Bretherton/Taylor
problem one can have difficulties with imposing the pressure boundary
conditions which do require to impose two separate boundary conditions for two
distribution functions.
 \item A number of simulations are done across the circular shape of channel,
which is quite difficult to address in terms of lattice Boltzmann free energy
framework where one needs to calculate the laplacians near the boundaries.
Although one can apply certain finite difference stencils
\cite{arnold-boundary,hunt-boundary} developed for the curved boundaries
however the analysis of error which can be addressed to the boundary conditions
treatment becomes an issue. 
\end{itemize}
Given all the concerns and challenges the suggested Lattice Boltzmann framework
is the two-dimensional flow driven by gravity or by pressure boundary
conditions. The advantage of the gravity driven bubble is that it's easy to
implement with the periodic boundary conditions. However, the analytical
solution to this problem, to the best authors' knowledge, is not yet known.
Though some authors \cite{sehgal-microchannel} use the gravity driven model to
validate the Bretherton film thickness, the latter should be used with caution
since the velocity of the bubble is different from the fluid velocity and there
is also the additional pressure drop due to front and rear menisci.

The solution
for relatively long bubbles driven by pressure specified on the inlet and the
outlet coincides with the classical theory of Bretherton. However, to implement
the pressure boundary conditions for the binary liquid model requires the
specification of boundaries conditions for two populations, for the momentum and
for the advection-diffusion equation (which are quite different). To the best
authors' knowledge such the situation is not yet covered in the literature. We
address this situation in Section \ref{appendix-pressure}.

\section{The bubble motion in the planar case}
This section covers the results for the pressure and force driven bubbles. The
idea is to establish what parameters are needed, i.e. grid size, bubble length,
viscosity ratio,to obtain the proper behaviour of Bretherton law.

\subsection{The flow driven by the gravity}
\citet{wong-films} indicated that the film width for polygonal capillaries
changes from $Ca^{1/2}$ on distances $x\propto Ca^{-1}$ to $Ca^{2}{3}$ for
distances $x \gg Ca^{-1}$. In simulations the bubble length is taken around 5-10
diameters of the channel {\color{red} Why is that? Is it because of
curvatures?}.

Also, gravity or force driven bubble flow has periodical boundary conditions.
Therefore, the situation to be simulated is the infinite train of bubbles
coming one after another. Thus, to avoid the influence of bubbles on each other
one needs to take quite long gap between bubbles. We take this gap around 3-5
lengths of the bubble.

\subsubsection{The influence of initial film width. Right parameters}
It is important to show that the steady film width is independent on initial
conditions where the most crucial condition is the initialized bubble width. If
in steady state the bubble width addjust itself to reveal certain film width
therefore the simulations are consistent. This section is to validate it.

For the sake of simplicity we picked up the following parameters for the
simulation:
\begin{equation}
\begin{aligned}
&Ca=0.005\\
&Ca=\frac{U_{bubble} \mu_{slug}}{\gamma}=0.005.
\end{aligned}
\end{equation}
Some parameters are fixed with the model parameters, as surface tension:
\begin{equation}
\gamma=\sqrt{\frac{8 k A}{9}}=\sqrt{\frac{8 0.04 0.04}{9}}=0.037712
\end{equation}

The film width for such small Capillary numbers is given by equation
\cite{giavedoni-numerical}:
\begin{equation}
\frac{h_{\inf}}{H}=1.3375 Ca^{2/3}=0.03899
\end{equation}
where we assume that $2 H$ or $Ny$ in lattice units is the channel width. That
means that one needs to resolve the interface which is around $2$ percent of
whole channel width with for instance $4$ lattice units. Therefore the channel
width is $200$ lattice units. We take the bubble length as $5$ channel widths
and the distance between slugs we take as $2$ bubble lengths. This will give
right away the grid size as $200\mathrm{x}3000$. Next step is to calculate the
velocity. The velocity can be calculated from the Capillary number as:
\begin{equation}
U_{bubble}=0.005 \frac{\gamma}{\mu_{slug}}=0.005 \frac{0.037712}{0.66666}=2.82\,
10^{-4},
\end{equation}
where $\mu_{slug}=1/3 (2.5-0.5)=0.6666$.
This velocity implies the following pressure gradient calculated from the
Poisuielle flow solution:
\begin{equation}
\begin{aligned}
&U_{bubble}=\frac{H^2}{8\mu} \frac{\mathrm{d}P}{\mathrm{d}x}=\frac{Ny^2}{8/3
(\tau_{liq}-1/2)}\frac{\mathrm{d}P}{\mathrm{d}x}\\
&\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{Ny^2}\gamma Ca
\end{aligned}
\end{equation}
Therefore the pressure gradient established is the following:
\begin{equation}
\begin{aligned}
\frac{\mathrm{d}P}{\mathrm{d}x}=3.77\, 10^{-8}.
\end{aligned}
\end{equation}

\subsubsection{The influence of initial film width. Easy parameters to check}
It is important to show that the steady film width is independent on initial
conditions where the most crucial condition is the initialized bubble width. If
in steady state the bubble width addjust itself to reveal certain film width
therefore the simulations are consistent. This section is to validate it.

The parameters in the previous subsection imply very long simulations. Even
Bretherton delivered his solution for the range of Capillary number less than
$0.005$. However, a lot of simulations devoted to the Bretherton problem
simulated extended it in the range of Capillary numbers greater than $0.005$
and for different Reynolds numbers. One can refer to figures
\ref{fig:giavedoni:planar} and \ref{fig:heil:planar} taken from citations
\cite{giavedoni-numerical,heil-bretherton} which crossvalidated their
simulations with Bretherton and each other. As far as this section is intended
to show physics behavior the moderate for simulation parameters are taken which
can be easily and fast validated on a computer. 
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/giavedoni_planar.eps}
\caption{\citet{giavedoni-numerical} gathered results across the
literature for different Capillary numbers \label{fig:giavedoni:planar}}
\includegraphics[width=0.97\textwidth]{Figures/heil-planar.eps}
\caption{\citet{heil-bretherton} performed simulations for different ranges of
Capillary numbers and Reynolds numbers. \label{fig:heil:planar}}
\end{figure}

One can see from the Fig. \ref{fig:heil:planar} that the width for the
Capillary number equal to $0.05$ ranges from $0.12$ to $0.15$ while Reynolds
number ranges from $0$ to $300$. This capillary number has moderate film widths
to be resolved with certain accuracy. Let us recalculate all the necessary
parameters:
\begin{equation}
\begin{aligned}
&Ca=0.05\\
&Ca=\frac{U_{bubble} \mu_{liq}}{\gamma}=0.05.
\end{aligned}
\end{equation}
Some parameters are fixed with the model parameters, as surface tension:
\begin{equation}
\gamma=\sqrt{\frac{8 k A}{9}}=\sqrt{\frac{8 0.04 0.04}{9}}=0.037712
\end{equation}
The film width for such small Capillary numbers is taken from
\ref{fig:heil:planar} for Reynolds number $40$ equals to $0.12$. We want to
resolve $12\%$ with $6$ lattice units. Therefore the channel
width is $50$ lattice units. We take the bubble length as $5$ channel widths
and the distance between slugs we take as $2$ bubble lengths. This will give
right away the grid size as $50\mathrm{x}750$. Next step is to calculate the
velocity. The velocity can be calculated from the Capillary number as:
\begin{equation}
U_{bubble}=0.05 \frac{\gamma}{\mu_{liq}}=0.05 \frac{0.037712}{0.6666}=2.82\,
10^{-3},
\end{equation}
where $\mu_{liq}=1/3 (2.5-0.5)=0.6666$. This implies the viscosity ratio to be
equal $R_{\nu}=2/0.2=10$ if we take the relaxation parameter for gas as
$\tau_{gas}=0.7$.
This velocity implies the following pressure gradient calculated from the
Poisuielle flow solution:
\begin{equation}
\begin{aligned}
&\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{Ny^2}\gamma Ca=6.003\,10^{-6}.
\end{aligned}
\end{equation}

Some pictures after $40000$ steps for the gravity driven flow and for different
initialization of bubble width as $Ny-12$, $Ny-20$, $Ny-28$.
\begin{figure}
\includegraphics*[bb=18 350 594 450,width=0.97\textwidth]
{Figures/force_width_6.eps}
\includegraphics*[bb=18 350 594 450,
width=0.97\textwidth]{Figures/force_width_10.eps}
\includegraphics*[bb=18 350 594 450,
width=0.97\textwidth]{Figures/force_width_14.eps}
\caption{The phase plots for different bubble widths as $Ny-12$,$Ny-20$ and
$Ny-28$.}
\end{figure}
\begin{figure}
\includegraphics[width=0.47\textwidth]{Figures/plane_force_width_6.eps}
\includegraphics[width=0.47\textwidth]{Figures/plane_force_width_10.eps}
\begin{center}
\includegraphics[width=0.47\textwidth]{Figures/plane_force_width_14.eps}
\end{center}
\caption{The phase profiles for different initialization parameters as the
bubble width $Ny-12$,$Ny-20$ and $Ny-28$.
\label{fig:phase:profiles:different:widths:force}}
\end{figure}

One can see from Figures \ref{fig:phase:profiles:different:widths:force} that
the film pattern in the beginning of the slug is almost independent of the
initial conditions and it ranges from 4 to 5 lattice boltzmann units, where the
is taken where the phase field has a value of zero. This preliminary result
justify that lattice Boltzmann method is able to predict the film thickness.

However, with the smaller initialization of the bubble one can see that the
front meniscus where the film thickness is validated accross is decreasing.
Therefore to be on the safe side one needs to initialize bubbles to be near the
analytical length.

We assumed that the interface thickness should be around $6$ lattice boltzmann
units, but we assume the calculations to be based on the $U_{slug}$ taken from
the simple Poisuille flow velocities. In reality the slug has slippage factor
with with the flow and velocity is slightly bigger than the liquid media. One
can take the velocity of the bubble from the simulations and recalculate all
the necessary lengths. Velocity profiles behind the bubble and in the front
meniscus can be seen in Fig. \ref{fig:velocity:profiles:different:widths:force}
and \ref{fig:velocity:profiles:different:widths:pressure} for the force driven
and pressure difference driven flows, correspondingly. As one can see average
velocity for the bubble and the slug larger than the calculated velocity which
should give larger the film. However, the film width for the planar case is
different and can have another coefficient in front of the proportionality to
$Ca^{2/3}$. Also, the phase gradient is taken as $0$ right now which can have
influence on the measurements as well.
\begin{figure}
\includegraphics[width=0.47\textwidth]{Figures/vel_bubble_force_width_6.eps}
\hfill
\includegraphics[width=0.47\textwidth]{Figures/vel_bulk_force_width_6.eps}\\
\includegraphics[width=0.47\textwidth]{Figures/vel_bubble_force_width_10.eps}
\hfill
\includegraphics[width=0.47\textwidth]{Figures/vel_bulk_force_width_10.eps}\\
\includegraphics[width=0.47\textwidth]{Figures/vel_bubble_force_width_14.eps}
\hfill
\includegraphics[width=0.47\textwidth]{Figures/vel_bulk_force_width_14.eps}\\
\caption{The velocity profiles in bulk and bubble for different initialization
parameters as the
bubble width $Ny-12$,$Ny-20$ and $Ny-28$.
\label{fig:velocity:profiles:different:widths:force}}
\end{figure}

While the shapes of droplets are different for the force and pressure
difference driven flows the velocity profiles show that the force driven bubble
move faster than in the case of pressure difference driven flow:
\begin{figure}
\includegraphics[width=0.47\textwidth]{Figures/vel_comparison_width_6.eps}
\hfill
\includegraphics[width=0.47\textwidth]{Figures/vel_comparison_width_10.eps}
\\
\begin{center}
\includegraphics[width=0.47\textwidth]{Figures/vel_comparison_width_14.eps}\\
\end{center}
\caption{Comparison between velocity bubble
profiles.\label{fig:velocity:comparison:different:widths}}
\end{figure}

Note that in the case of pressure driven flows the termination bulk and bubble
velocity have more consistent behavior and the values are not different too
much. That is one of the reasons why in the case of the force driven flows the
width has a bit different values depending on the initialization procedure.

>>>>

In terms of D2Q9 the equilibrium populations are as \cite{pooley-contact}:
\begin{equation}
\begin{aligned}
&f_i^{eq}=w_i 
\biggl(3
p_0 - k \phi \Delta \phi
+\frac{u_{\alpha}c_{i\alpha}}{c_s^2}+\frac{Q_{i\alpha\beta}u_{\alpha } u_ {
\beta}}{2 c_s^4}\biggr), i=1\div8\\
&f_0^{eq}=\rho-\sum_{i\neq0}{f_i^{eq}}\\
&g_i^{eq}=w_i(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi^m
\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}), i=1\div8 \\
&g_0^{eq}=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
\end{aligned}
\end{equation}
where $\Gamma$ is the mobility parameter; chemical potential
$\mu=-A\phi+A\phi^3-k\Delta\phi$; $k$ is related to the surface
tension; $A$ is the parameter of the free-energy model; bulk pressure
$p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$. D2Q9 related parameters are
$w_i=\left\{\frac{4}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},\frac{1}{9},
\frac{1}{36},\frac{1}{36},\frac{1}{36},\frac{1}{36}\right\}$, and
$Q_{i\alpha\beta}=c_{i\alpha} c_{i\beta} - c_s^2 \delta_{\alpha\beta}$ with
the sound speed parameter $c_s^2=\dfrac{1}{3}$. Weights related to the
inclusion of the surface tension coefficient are as follows:
$w^{xx}_{1-2}=w^{yy}_{3-4}=1/3$, $w^{xx}_{3-4}=w^{yy}_{1-2}=-1/6$,
$w^{xx}_{5-8}=w^{yy}_{5-8}=-1/24$, $w^{xy}_{1-4}=0$, $w^{xy}_{5-6}=1/4$ and
$w^{xy}_{7-8}=-1/4$. Note that the system restores the macroscopic
equations as:
\begin{equation}
\begin{aligned}
&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}=
-\partial_{\alpha}P_{\alpha \beta} +
\nu\partial_{\beta}\left(\partial_{\alpha}u_{\beta}+\partial_{\beta} u_{\alpha}
+ \frac{1}{3}\partial_{\gamma} u_{\gamma} \delta_{\alpha\beta}\right)\\
&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \Delta \mu,
\end{aligned}
\label{binary:fluid:system}
\end{equation}
where the viscosity $\nu=c_s^2 (\tau-1/2)$ and mobility parameter
$M=\Gamma(\tau_{\phi}-1/2)$. The system allows the separation of the liquid
phase with $\phi=1$ and so-called gas phase with $\phi=-1$. Note that the
spatial variation of the relaxation parameter depends on the relaxation
parameter of $\tau_{gas}$ and $\tau_{liq}$ linearly as
$\tau=\tau_{gas}+\frac{\phi+1}{2}\tau_{liq}$. Note that the surface tension in
the framework of binary liquid model as $\sqrt{8 k A}{9}$.

\section{Lattice Botlzmann benchmark}
To properly simulate and to compare simulations results with the already
performed simulations one needs to design a numerical benchmark by addressing
the following challenges for the lattice Boltzmann method:
\begin{itemize}
 \item The diffusive interface nature of the multiphase model should not
influence the film width results. This is obviously achieved only for certain
resolutions of the diffuse interface in comparison with the film width.
 \item The spurious currents arising from the surface tension discretization
can influence the mass transfer in the thin film therefore comprimising the
overall solution. Therefore the bubble should be long enough to decrease the
influence of the spurious currents.
 \item The wettability coefficient which is in principle can control the
dynamic angle and menisci appearance \cite{pagonabarraga-finger} should not
influence on the overall solution.
 \item The free-energy model examined in the work is one density/different
viscosities model. Classical Bretherton problem is the free-interface problem
where the film thickness is established by itself. The viscosities ratio should
be large as it is indicated in \cite{pagonabarraga-fingers}, however no studies
address this issue.
  \item The Bretherton problem addresses the bubbles flow driven by pressure
difference but not body force. However, it's relatively difficult to impose the
pressure difference in the framework of the binary liquid model because two
boundary conditions with the associated gradients and laplacians are required.
The pressure boundary conditions for the binary liquid model are currently
under development. Preliminary results show that the differences between
pressure difference flows and body force driven flows are small for the film
deposition thickness, the interest of this study. Thus, the work's focus is at
the body force driven flows because of the better simulation stability and
realization along with the periodic boundary conditions.
 \item Most simulations are done across the circular shape of channel,
which is quite difficult to address in terms of lattice Boltzmann free energy
framework where one needs to calculate the laplacians near the boundaries.
Although one can apply certain finite difference stencils
\cite{arnold-boundary,hunt-boundary} developed for the curved boundaries
however the analysis of error which can be addressed to the boundary conditions
treatment becomes an issue. Therefore the benchmark is for an axis-symmetric
case. 
\end{itemize}
Given all the concerns and challenges the suggested Lattice Boltzmann framework
is the two-dimensional flow driven by a body force. However, the analytical
solution to this problem, to the best authors' knowledge, is not yet known.
Though some authors \cite{sehgal-microchannel} use the gravity driven model to
validate the Bretherton film thickness, the latter should be used with caution
since the velocity of the bubble is different from the fluid velocity and a
bubble shape
is different in front and rear menisci. \citet{wong-films} indicated that the
film width for polygonal capillaries
changes from $Ca^{1/2}$ on distances $x\propto Ca^{-1}$ to $Ca^{2/3}$ for
distances $x \gg Ca^{-1}$. Therefore, for a bubble to have a formed shape in
the  front meniscus we picked up the bubble length as $5$ channel heights. The
channel length is taken to minimize the influence of one bubble on another,
because of the periodicity of boundary conditions. For the simulations the
channel length is taken as $3$ bubble lenghts giving all together $15$ channel
heights. Figure \ref{fig:benchmark:sketch} represents the benchmark sketch.
\begin{figure}
%\includegraphics[width=0.97\textwidth]{}
\caption{{\color{red} Ask Sasha to help here. The benchmark sketch.}
\label{fig:benchmark:sketch}}
\end{figure}

\section{Results}
\subsection{The nondimensionalization and initialization procedure}
This section is to introduce the procedure of nondimensionalization and
initialization techniques to be able to predict parameters necessary for
simulation. The capillary number governs the interface thickness and defined as:
\begin{equation}
Ca=\frac{U_{bubble} \mu_{liq}}{\gamma},
\end{equation}
where $\mu_{liq}$ is the liquid viscosity, $\gamma$ is the surface tension. The
bubble velocity $U_{bubble}$ is taken in the center of the bubble where the film
thickness is examined. 
While parameters as $\mu_{liq}$, $\gamma$ and capillary number $Ca$ are given
or defined through the model parameters, the velocity $U_{bubble}$ is not
defined. In the simulation we approximately match the velocity with the
Poiseuille profile center velocit, which is given by the equation:
\begin{equation}
\begin{aligned}
U_{bubble} \approx U_{liq}=\frac{H^2}{8
\mu_{liq}}\frac{\mathrm{d}P}{\mathrm{d}x}\\
U_{bubble}\approx
\frac{{N_y}^2}{\frac{8}{6}(\tau_{liq}-\frac{1}{2})}\frac{\mathrm{d}P}{\mathrm{d}
x } ,
\end{aligned}
\end{equation}
By taking the Poiseuille profile one can estimate the body force through the
capillary number and grid resolution:
\begin{equation}
\begin{aligned}
&\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{{N_y}^2}\gamma Ca.
\end{aligned}
\end{equation}
In reality we iterate the body force value
until the desired velocity and the capillary number as well are achieved.

After one resolves necessary parameters to run a simulation, one needs to
initialize the macroscopic fields and populations.
The velocity is initialized with $0$ everywhere. Populations are initialized
with the equilbrium values for the binary liquid model including all the phase
gradients and laplacians. The liquid phase is initialized with $1$ and the
gas phase is initialized with $-1$. Though the initialization film thickness
is not involved in the final result for the front meniscus, Fig.
\ref{fig:different:initialization:widths}, we keep the initialized film
thickness as close as possible to the already obtained numerical film
thicknesses \cite{giavedoni-numerical}. Otherwise, the region where the
Bretherton phenomena occurs is dillute, which is obviously seen for the
simulation initialized with the film thickness as $Ny-28$ {\color{red} Michal,
can you please prepare the simulations with the different initialization. Or
just delete different thicknesses initialization}.
\begin{figure}
\includegraphics*[bb=18 350 594 450,width=0.97\textwidth]
{Figures/force_width_6.eps}
\includegraphics*[bb=18 350 594 450,
width=0.97\textwidth]{Figures/force_width_10.eps}
\includegraphics*[bb=18 350 594 450,
width=0.97\textwidth]{Figures/force_width_14.eps}
\caption{\color{red} The phase plots for different bubble widths as
$Ny-12$,$Ny-20$ and
$Ny-28$. Michal, could you please prepare nice pictures for
simulations. Indicate as well the
grid and time steps\label{fig:different:initialization:widths}}
\end{figure}

Here are two different examples presented that can quantitatively show the
initialization procedure for quite different capillary numbers. In these
examples we use $k=0.04$ and $A=0.04$ as optimal binary liquid model parameters.
That explicitly implies the surface tension to be $\sqrt{\frac{8 k
A}{9}}=0.0377$ and the characteristic length of the interface
$4\sqrt{\frac{k}{A}}=4$ lattice boltzmann units. For the sake of simplicity and
stability the relaxations parameter for liquid and gas phases are taken as
$\tau_{liq}=2.5$ and $\tau_{gas}=0.7$. That gives the gas-liquid viscosity ratio
as $10$.  
\begin{description}
 \item[I \bm{$Ca=0.005$}]
  The film width for such small Capillary numbers \cite{bretherton} is
given by equation:
  \begin{equation}
  \frac{h_{\infty}}{H}=1.3375 Ca^{2/3}=0.03899,
  \end{equation}
  where we assume that $2 H$ or $Ny$ in lattice units is the channel width.
Therefore one needs to resolve the interface which is around $2$ percent of
  whole channel width with for instance $8$ lattice units, where $4$ lattice
units occupies the interface width for given. Note that the interface occupies
$50$ percent of the interface.  Therefore the channel
  width is $400$ lattice units. We take the bubble length as $5$ channel widths
  and the distance between bubble is as $2$ bubble lengths. This will give
  right away the grid size as $400\mathrm{x}6000$. In terms of computer memory
requirement the calculations with the double precision accuracy demand:
\begin{equation}
\begin{aligned}
&2\text{ set of populations }\mathrm{x}\,\, 8\,\, \text{bytes per population
}\mathrm{x}\,\, 9 \text{ number of
populations}\\
&\mathrm{x}\,400\,\mathrm{x}\,6000\approx 345 \mathrm{MB}. 
\end{aligned}
\end{equation}
Even with this naive assumption for the $50$ percent resolution of the film
widht in comparison with the interface width, one can get the memory
requirement as $345 MB$.

After resolving the interface one can  calculate the
  velocity from the capillary number as:
  \begin{equation}
  U_{bubble}=0.005 \frac{\gamma}{\mu_{slug}}=0.005
\frac{0.0377}{0.66666}=2.82\,10^{-4},
  \end{equation}
  where $\mu_{slug}=1/3 (2.5-0.5)=0.6666$.
  This velocity implies the following pressure gradient calculated from the
  Poisuielle flow solution:
  \begin{equation}
  \begin{aligned}
  &U_{bubble}=\frac{H^2}{8\mu} \frac{\mathrm{d}P}{\mathrm{d}x}=\frac{Ny^2}{8/3
  (\tau_{liq}-1/2)}\frac{\mathrm{d}P}{\mathrm{d}x}\\
  &\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{Ny^2}\gamma Ca
  \end{aligned}
  \end{equation}
  Therefore the pressure gradient established is the following:
  \begin{equation}
  \begin{aligned}
  \frac{\mathrm{d}P}{\mathrm{d}x}=9.42\, 10^{-9}.
  \end{aligned}
  \end{equation}
  Note that in the case of large grids  height resolution is directly
involved in	
the physical time step as:
\begin{equation}
\Delta t =\frac{U_{LBM}}{U_{phys}} \Delta x ,
\end{equation}
where for example the real velocities in the physical world can reach
approximately $1$ m/s and physical channel height upto $1$ mm. In this case one
can easily approximate the physical time step:
\begin{equation}
\Delta t = 7.05 \, 10^{-10} \,\mathrm{s}
\end{equation}
which is a small time step. The typical simulation needs approximately
$100,000$ multiplication time steps to come to the steady state solution.

 \item[II $\bm{Ca=0.05}$] 
   The parameters in the previous subsection imply very long
simulations. Bretherton delivered his solution for the range of Capillary number
less than
  $0.005$. However, a lot of simulations devoted to the Bretherton/Taylor
flow
  simulated and extended it in the range of Capillary numbers greater than
$0.005$
  and for different Reynolds numbers. One can refer to figures
  \ref{fig:giavedoni:planar} and \ref{fig:heil:planar} taken from citations
  \cite{giavedoni-numerical,heil-bretherton} which crossvalidated their
  simulations with Bretherton and each other. Even though LBM can simulate the
flows with small capillary numbers, i.e. less than $0.005$, but waiting time
for the simulation to be accomplished is relatively long. Therefore, we focus
our work on the simulations of \citet{giavedoni-numerical} which crossvalidated
results not only for small capillary numbers but for large capillary numbers as
well. As far as the work intention is
  to show proper physics behavior the moderate parameters are
taken which
  can be easily and fast validated on a computer. 
  One can see from the Fig. \ref{fig:heil:planar} that the width for the
  Capillary number equal to $0.05$ ranges from $0.12$ to $0.15$ while Reynolds
  number ranges from $0$ to $300$. This capillary number has moderate film
widths
  to be resolved with certain accuracy.

  The film width for such small Capillary numbers is taken from
  \ref{fig:heil:planar} for Reynolds number $40$ equals to $0.12$. We want to
  resolve $6\%$ of whole channel width with $12$ lattice units, where $4$
lattice units are related to the interface resolution.  Therefore the channel
  width is $200$ lattice units. We take the bubble length as $5$ channel widths
  and the distance between slugs we take as $2$ bubble lengths. This will give
  right away the grid size as $200\mathrm{x}3000$. The velocity can be
calculated from the Capillary number as:
  \begin{equation}
  U_{bubble}=0.05 \frac{\gamma}{\mu_{liq}}=0.05 \frac{0.037712}{0.6666}=2.82\,
  10^{-3}.
  \end{equation}
  The velocity implies the following pressure gradient calculated from the
  Poisuielle flow solution:
  \begin{equation}
  \begin{aligned}
  &\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{Ny^2}\gamma Ca=3.77\,10^{-7}.
  \end{aligned}
  \end{equation}
  The initialization approach is to take values from the already performed
numerical simulation and to calculate the grid numbers from them matching the
body force gradient through the solution of Poiseuille flow.

\end{description}
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/giavedoni_planar.eps}
\caption{\citet{giavedoni-numerical} gathered results across the
literature for different Capillary numbers \label{fig:giavedoni:planar}}
\includegraphics[width=0.97\textwidth]{Figures/heil-planar.eps}
\caption{\citet{heil-bretherton} performed simulations for different ranges of
Capillary numbers and Reynolds numbers. \label{fig:heil:planar}}
\end{figure}
Note that we base the initialization techniques through correlations for the
capillary number and the body force. However, in the binary liquid framework
those correlations give only the approximation. First of all the body force
which is based on the Poiesuille velocity profile for the liquid phase is the
biggest assumption. In reality the less viscous bubble will have a slippage
factor and its velocity is larger than the surrounding liquid velocity. By
plugging in the obtained velocity back to the capillary number equation the
larger capillary number can be obtained. While we present the initialization
tips in reality one needs to take the bubble velocity from the simulations and
recalculate all the necessary characteristics.

\subsubsection{The influence of initial film width. Easy parameters to check}

 Velocity profiles behind the bubble and in the front
  meniscus can be seen in Fig.
\ref{fig:velocity:profiles:different:widths:force}
  and \ref{fig:velocity:profiles:different:widths:pressure} for the force driven
  and pressure difference driven flows, correspondingly. As one can see average
  velocity for the bubble and the slug larger than the calculated velocity which
  should give larger the film. However, the film width for the planar case is
  different and can have another coefficient in front of the proportionality to
  $Ca^{2/3}$. Also, the phase gradient is taken as $0$ right now which can have
  influence on the measurements as well.
  \begin{figure}
  \includegraphics[width=0.47\textwidth]{Figures/vel_bubble_force_width_6.eps}
  \hfill
  \includegraphics[width=0.47\textwidth]{Figures/vel_bulk_force_width_6.eps}\\
  \includegraphics[width=0.47\textwidth]{Figures/vel_bubble_force_width_10.eps}
  \hfill
  \includegraphics[width=0.47\textwidth]{Figures/vel_bulk_force_width_10.eps}\\
  \includegraphics[width=0.47\textwidth]{Figures/vel_bubble_force_width_14.eps}
  \hfill
  \includegraphics[width=0.47\textwidth]{Figures/vel_bulk_force_width_14.eps}\\
  \caption{The velocity profiles in bulk and bubble for different initialization
  parameters as the
  bubble width $Ny-12$,$Ny-20$ and $Ny-28$.
  \label{fig:velocity:profiles:different:widths:force}}
  \end{figure}
